#####BAX452 HW4#####
```{r}
library(parallel)
library(ggplot2)
library(GGally)
library(FNN)
library(mvtnorm)
```
### Exercise 1.22
### 1 Try doing several runs of the linear and k-NN code in that section, comparing results.

```{r}

# linear model
library(freqparcoord)
data(mlb)

xvalpart <- function(data, p) {
  n <- nrow(data)
  ntrain <- round(p*n)
  trainidxs <- sample(1:n, ntrain, replace = FALSE)
  list(train = data[trainidxs,], valid = data[-trainidxs,])
}

xvallm <- function(data, ycol, predvars, p, meanabs = TRUE){
  tmp <- xvalpart(data, p)
  train <- tmp$train
  valid <- tmp$valid
  trainy <- train[, ycol]
  trainpreds <- train[, predvars]
  trainpreds <- as.matrix(trainpreds)
  lmout <- lm(trainy ~ trainpreds)
  validpreds <- as.matrix(valid[, predvars])
  predy <- cbind(1, validpreds) %*% coef(lmout)
  realy <- valid[, ycol]
  if (meanabs) return(mean(abs(predy - realy)))
  list(predy = predy, realy = realy)
}


xvallm(mlb,5 ,c(4, 6), 2/3)
```
# run 5 times, results are different
#[1] 14.53308
#[1] 14.10442
#[1] 13.39469
#[1] 13.46165
#[1] 13.7847



```{r}
#k-NN model

library(dummies)
library(regtools)
set.seed(9999)

xvalknn <- function(data, ycol, predvars,k, p, meanabs = TRUE){
  data <- data[, c(predvars, ycol)]
  ycol <- length(predvars) + 1
  tmp <- xvalpart(data, p)
  train <- tmp$train
  valid <- tmp$valid
  valid <- as.matrix(valid)
  xd <- preprocessx(train[, -ycol],k)
  kout <- knnest(train[,ycol], xd,k)
  predy <- predict(kout, valid[,-ycol], TRUE)
  realy <- valid[, ycol]
  if (meanabs) return(mean(abs(predy - realy)))
  list(predy = predy, realy = realy)
}

xvalknn(mlb,5, c(4,6), 25, 2/3)
```
# run 5 times, results are different
#[1] 14.32817
#[1] 14.13207
#[1] 13.73988
#[1] 14.16343
#[1] 14.1968
#The two methods gave similar results.


###2 . Extend (1.28) to include interaction terms for age and gender, and age2 and gender. 
### Run the new model, and find the estimated effect of being female, for a 32-year-old person with a Master's degree.
```{r}
library(freqparcoord)
data(prgeng)
prgeng$age2 <- prgeng$age^2
edu <- prgeng$educ
prgeng$ms <- as.integer(edu == 14)
prgeng$phd <- as.integer(edu == 16)
prgeng$fem <- prgeng$sex - 1
tmp <- prgeng[edu >= 13,]
pe <- tmp[, c(1,12,9,13,14,15,8)]
pe <- as.matrix(pe)

#adding age and gender, age2 and gender
lm_wage = lm(wageinc ~ age + age2 + wkswrkd + ms + phd + fem + 
     ms:fem + phd:fem + age:fem + age2 : fem, data = prgeng)
summary(lm_wage)
```

```{r}
predict(lm_wage, data.frame(age = 32,age2 = 32^2,wkswrkd = 52, ms = 1, phd = 0 ,fem = 1))

```

# 32 yrs old female with MS degree earns $67132 wage income


### 3 Use lm() to form a prediction equation for density from the other variables (skipping the first three)
### comment on whether use of indirect methods in this way seems feasible.
```{r}
bodyfat <- read.csv("bodyfat.csv")
lm_density = lm(density ~ age + weight + height + neck + chest + abdomen + hip + thigh + knee + ankle + biceps + forearm + wrist, data = bodyfat)
summary(lm_density)
```
# from the summary we can see there are several variable are significat (p-value <0.05) predictors of density
# indirect method seems a feasible way since abdomen, wrist, forearm and neck are easier to measure.


### 4 

#(a) Write English prose that relates the overall mean height of people 
# and the gender-specific mean heights.

# The overall mean height of people is a weighted average of the gender-specifiec mean heights,
# with the weight for each gender being its proportion of the overall population.

#(b) Write English prose that relates the overall proportion of people 
# taller than 70 inches to the gender-specific proportions.

# The overall proportion of people taller than 70 inches is composed by 
# proportion of female taller than 70 inches and proportion of male taller than 70 inches



### Exercise 2.14

### 1 Consider the census data

```{r}
### (a) Form an approximate 95% confidence interval for b6 in the model

library(freqparcoord)
data(prgeng)
prgeng$age2 <- prgeng$age^2
edu <- prgeng$educ
prgeng$ms <- as.integer(edu == 14)
prgeng$phd <- as.integer(edu == 16)
prgeng$fem <- prgeng$sex - 1
tmp <- prgeng[edu >= 13,]
pe <- tmp[, c(1,12,9,13,14,15,8)]
pe <- as.matrix(pe)

lm2_wage = lm(wageinc ~ age + age2 + wkswrkd + ms + phd + fem, data = prgeng)
summary(lm2_wage)

```

```{r}
# CI for b6
b6_l95 = round((-11484.49 - 1.96 * 705.30), 2)
b6_r95 = round((-11484.49 + 1.96 * 705.30), 2)
sprintf("95 percent Confidence Interval for B6 female: (%s, %s)", b6_l95, b6_r95)
```

```{r}
### (b) Form an approximate 95% confidence interval for the gender effect
### for Master's degree holders, b6 + b7, in the model.
lm3_wage = lm(wageinc ~ age + age2 + wkswrkd + ms + phd + fem + 
                +                ms:fem + phd:fem , data = prgeng)
summary(lm3_wage)

```

```{r}
#CI for b6+b7
b6b7_l95 = round((-4157.253 - 1.96 * 1728.329), 2)
b6b7_r95 = round((-4157.253 + 1.96 * 1728.329), 2)
sprintf("95 percent Confidence Interval for B6+B7: (%s, %s)", b6b7_l95, b6b7_r95)
```
### 2 Extend the analysis in Section 2.8.5 to the full data set, 
### adding dummy variables indicating the second and third year. 
### Form an approximate 95% confidence interval for the difference between the coefficients of these two dummies.
```{r}

shar <- read.csv('day.csv')
shar$temp2 <- shar$temp^2
shar$clearday <- as.integer(shar$weathersit == 1)
# yr=1 indicates 2nd yr, there's no data for 3rd yr
lm_shar <- lm(registered ~ temp + temp2 + workingday + clearday + yr, data = shar)
summary(lm_shar)

```

```{r}
yr_l95 = round((1716.25 - 1.96 *  56.68), 2)
yr_r95 = round((1716.25 + 1.96 *  56.68), 2)
sprintf("95 percent Confidence Interval for B6+B7: (%s, %s)", yr_l95, yr_r95)
```

###3 Explain why each Di is (k-1)-variate normal, 
### and derive matrix expressions for the mean vector and covariance matrices.

# Since Hi is k-variate normal and Dij = Hi,j+1 - Hij, j=1,2,...k-1
# at age k-1, Di,k = Hi,k - Hi,k-1, there's no Di at age K
# Di is (k-1) variate normal
# mean vector is u = E[Di] = [E[D1], E[D2],...,E[Dk-1]]^T
# coveriance matrices is Sum = E[(Di- u)(Di - u) ^T] = [Cov[Di,Dj]; 1<= i, j<= k-1]


### 4 In the simulation in Section 2.9.3, it is claimed that p2 = 0:50. 
### Confirm this through derivation.

# set variance of xj =1 Y= x1+..+Xp and variance of e = p
# p^2 = 1 - Var(e) / Var(Y) 
#     = 1 - Var(e) / (Var(u(X)) + Var(e)) 
#     = 1 - p/(1*p+p) = 0.5